---
title: "Comparison of the exponential distribution and the Central Limit Theorem "
author: "Cindy Chen"
date: "Oct 20, 2015"
output: pdf_document
---
```{r echo = FALSE, message=FALSE, warning=FALSE}
```
# Statistical Inference Project

## Overview:   

In this project we will investigate the exponential distribution in R and compare 
it with the Central Limit Theorem(CLT).  

The exponential distribution can be simulated in R with 'rexp(n, $\lambda$)' where 
lambda $\lambda$ is the rate parameter. 

The mean of exponential distribution ($\mu$) is 1/$\lambda$ and the standard 
deviation ($\sigma$) is also 1/$\lambda$. 

According the CLT if a sample consists of at least 30 independent observations 
and the data are not strongly skewed, then the distribution of the sample mean 
is approximattely:  $\bar{x}_{n}$ ~ $N(\mu, \frac{\sigma }{\sqrt{n}})$.

In this analysis, we will show that the sampling distribution of the mean of an 
exponential distribution with $n$ = 40 observations and $\lambda$ = 0.2. The $Exp(\frac{1}{0.2}, \frac{1}{0.2})$ distribution then is approximately 
$N(\frac{1}{0.2}, \frac{\frac{1}{0.2}}{\sqrt{40}})$.

We will investigate the distribution of averages of 40 exponentials and we will 
need to do 1000 simulations.


## Simulations: 
Here is the R code for 1000 simulations:
```{r}
set.seed(123)
# The distribution of 1000 averages of size 40 from the exponential distribution
exp_sample_means <- NULL
for(i in 1:1000) {
  exp_sample_means <- c(exp_sample_means, mean(rexp(40, 0.2)))
}
```  


## Sample Mean versus Theoretical Mean:

In the following we will draw 1000 samples of size 40 from an $Exp(\frac{1}{0.2}, \frac{1}{0.2})$ distribution. For each of the 1000 samples we will calculate the mean.
(Then we will get 1000 means)

Theoretically, this is the same as drawing a single sample of size 1000 from the corresponding sampling distribution with $N(\frac{1}{0.2}, \frac{\frac{1}{0.2}}{\sqrt{40}})$.

$\bar{x}$ in our case is `r mean(exp_sample_means)` which is very close 
to the mean of the theoretical distribution namely $\mu = \frac{1}{0.2} = 5$.

##Sample Variance versus Theoretical Variance:   

According to the CLT we would expect that the variance of the sample of the 1000 
means is approximately $\frac{\frac{1}{0.2^{2}}}{40} = 0.625$.
```{r}
var(exp_sample_means)
```
$s^{2}$ in our case is `r var(exp_sample_means)`, which is close to the 
variance of the theoretical distribution 0.625.


## Showing the sample mean distribution is approximately normal:  

In order to demonstrate that the sample distribution of the 1000 means is 
approximately normal, we will plot the histogram of the distribution of the sample
mean $\bar{x}_{n}$, and overlay it with the density function from the theoretical 
sampling distribution which is approximately $N(\frac{1}{0.2}. \frac{\frac{1}{0.2}}{\sqrt{40}})$ 

```{r}
library(ggplot2)
data <- as.data.frame(exp_sample_means)
ggplot(data, aes(x = exp_sample_means)) + 
  geom_histogram(binwidth = 0.3, color = 'black', fill = 'steelblue', aes(y = ..density..))+
  stat_function(aes(x = c(2, 8)), fun = dnorm, color = 'red', 
                args = list(mean = 5, sd = sqrt(0.625))) +
  xlab('Sample mean') +
  ylab('Density') +
  ggtitle('Comparison of the distribution of sample mean\n and the theoretical distribution')
```


# Conclusions
In this analysis, we showed that the sampling distribution of the mean of an exponential distribution with $n = 40$ observations and $\lambda = 0.2$ is approximately $N(\frac{1}{0.2}, \frac{\frac{1}{0.2}}{\sqrt{40}})$.

Thus, we showed:  
1. The sample mean is close to the theoretical mean of the distribution.  
2. The variable of the sample mean distribution is close to the theoretical 
variance of the distribution.  
3. The sample mean distribution is approximately normal.  
